Comparing MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration

### 2023-10-22 manuscript/main.tex
Final Submission

\begin{abstract}
    Simulation based calibration (SBC) \citep{talts2020validating} is applied to analyse two commonly used, competing Markov chain Monte Carlo algorithms for estimating the posterior distribution of a stochastic volatility model. In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}. The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through effective sample size. Results show that Stan's No-U-Turn sampler, an implementation of Hamiltonian Monte Carlo, produces a well-calibrated posterior estimate while the celebrated off-set mixture approach is less efficient and poorly calibrated, though model parameterisation also plays a role. Limitations and restrictions of generality are discussed.

@@

@@

    This research analyses and compares the calibration and efficiency of MCMC used to estimate stochastic volatility models using a simulation study. By calibration we refer to a MCMC algorithm that returns correct posterior, on average, conditional on the correct model specification. The first algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}. The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm. The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates on average.

    Comparisons of these algorithms are conducted using a methodology known as Simulation Based Calibration (SBC) \citep{talts2020validating}. Parameters are drawn from the prior distribution and used to create data sets from the generative stochastic volatility model. We then obtain posterior samples from each of the MCMC methods. Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters given the data generating process. The key diagnostic metrics are the effective sample size, to measure the efficiency of the algorithm, and uniform rank statistics to determine calibration of the posterior estimates.

    One particular challenge to comparing models is the parameterisation. Different parameterisations of a model, whilst mathematically the same, can be very different in terms of sampling difficulty \citep{neal2003slice}. In this work two different parameterisations of the stochastic volatility model are compared for each sampling method. An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.

@@

\subsection{Challenge 1: Comparing posteriors with no ground truth}
    The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms. Convergence metrics assessed from the same MCMC output to be used for posterior inference are commonly used to check the performance of the MCMC chains. Most often, the effective sample for each parameter is produced, as it indicates the number of approximately independent draws obtained from the dependent Markov chain for a target parameter \citep{gelman2013bayesian}, and $\hat{R}$, the potential scale reduction factor or measure of variability between different chains \citep{gelman1992inference}, is computed, to gauge whether the MCMC algorithm has converged to the target posterior distribution.

@@

\subsection{Challenge 2: Limitations of a single simulation}
    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not observe the true data generating process or the true model. A strategy around this is to evaluate a model and algorithm on simulated data. One approach is to simulate data from a generative model using known parameters. Then fit the same model on the simulated data and see if the true parameters can be recovered. This gives us the benefit of defining the true parameters of the data generating process to be estimated. If the model and algorithm cannot adequately capture the true parameter, then we cannot be confident that it will provide reliable estimates on real data.

@@

    Figure 2 shows marginal posterior distributions given a single set of 1000 returns generated by the SV model. The true parameters are $\mu=-0.389$, $\sigma^2_{\eta}=0.0115$ and $\phi=0.547$ and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler. The 95\% credible intervals for $\mu$ and $\sigma^2_{\eta}$ cover the true parameter. The true parameter for $\phi$ however, is in the tails and outside the interval. Such an analysis may incorrectly conclude that the model fails to adequately estimate the $\phi$ parameter. However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated data set.

@@

\subsection{Research Goal}
    The objective of this research is to design a simulation study to evaluate the calibration of the HMC and KSC off-set mixture algorithms used to estimate SV models. As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations. To check the calibration of an algorithm, repeated independent simulations are required. This methodology is discussed in the next section.

@@

\section{Methodology}
This section introduces the simulation design used to evaluate whether a MCMC algorithm is returning the correct posterior estimates on average (that is, whether a MCMC algorithm calibrated). Section 3.1 discusses the procedure for running simulation based calibration and Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.\footnote{The software used to implement these methods and produce all results and plots in this paper are the R v4.2.3\citep{rlang}, Python v3.8.13 \citep{10.5555/1593511} and Stan v2.31.0 \citep{stan} programming languages, including open source R packages (tidyverse v1.3.2 \citep{tidyverse}, quantmod v0.4.20 \citep{quantmod}, cmdstanr 0.5.3 \citep{cmdstanr}, posterior v1.4.1 \citep{posteriorr}, bayesplot v1.4.10 \citep{bayesplot}, jsonlite v1.8.4 \citep{jsonlite}, arrow v12.0.1.1 \citep{arrow} and future.apply v1.11.0 \citep{RJ-2021-048}) and python packages (numpy v1.24.4 \citep{harris2020array}, scipy v.1.10.1 \citep{2020SciPy-NMeth}, statsmodels v.0.14.0 \citep{seabold2010statsmodels}, pandas v.2.0.3 \citep{mckinney-proc-scipy-2010} and pyarrow v12.0.1 \citep{arrow}).}

    \subsection{Simulation Design}
        Simulation Based Calibration (SBC) checks the calibration of posterior estimates generated by MCMC algorithms. SBC is conducted by comparing the distribution of rank statistics to the uniform distribution which arises when an algorithm is correctly calibrated. For each $k$ of $K$ SBC iterations, one value for each parameter is drawn from the prior. Conditional on the drawn parameter values, a data set is generated from the model. Then a sampling method (originally using Stan in \citet{talts2020validating}) is used to obtain posterior samples given this data and the prior. The estimated posteriors are then compared to the true values by calculating the rank statistics. Under a calibrated model, the $K$ rank statistics for each parameter should be uniformly distributed.

        To illustrate the SBC procedure, let $\theta$ be an arbitrary scalar parameter from a model and $y$ represent a data set. Start with a single parameter $\theta^{sim}$ drawn from the prior distribution
        \begin{align}
        \theta^{sim} \sim \pi(\theta),
        \end{align}
        then generate a data set $y^{sim}$ conditionally given $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, notation to condition on the model is suppressed.}:
        \begin{align}
        y^{sim} \sim \pi (y|\theta^{sim}).
        \end{align}
        Take $L$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC):
        \begin{align}
        \{\theta_1,\dots , \theta_{B}\} \sim \pi (\theta | y^{sim}).
        \end{align}

        \citet{talts2020validating} prove that the sample of rank statistics for a given parameter, produced from replicate posteriors, follow a uniform distribution over the integers $\{0,1\dots,B\}$ if the posterior samples follow the same distribution as the prior on average. The rank statistic for any one-dimensional function of parameters $f:\Theta\rightarrow\ \mathbb{R}$ produced by the $k^{\mathrm{th}}$ replicate posterior sample, is defined as
        \begin{align}
        r^{(k)} = \mathrm{rank}(\{f(\theta_1^{(k)}),\dots , f(\theta_{B}^{(k)})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}^{(k)}) < f(\theta^{sim})].
        \end{align}

        This completes one iteration of SBC. To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter. The resulting rank statistics are compared to the uniform distribution to determine if the algorithm is calibrated and returning the correct posteriors.

        A key result is that the posterior averaged over the data and true parameters will equal the prior distribution. This is shown in the following expression
        \begin{align}
        \pi(\theta) &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})d\theta^{sim} dy^{sim} \\
        &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) d\theta^{sim} dy^{sim}.
        \end{align}
        That is, the average posterior for $\theta$ over the SBC iterations should approximately equal to prior distribution, if the algorithm is calibrated. Therefore, any deviation between the prior distribution and the average posterior draws produced from replicated posterior distributions means that the sampling methodology is not producing the correct posterior, on average.

@@

            The shape of the distribution of rank statistics gives insight into how an MCMC may be miscalibrated. Specifically, it gives information about how the average posterior deviates from the prior distribution of a given parameter. Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2020validating} for convenience. These display examples of posterior miscalibration and the corresponding non-uniform rank statistics.

            Figure \ref{fig:underestimation} shows a distribution of rank statistics with inflated frequency on the right side of the histogram. This distribution is consistent with posterior samples that tend to underestimate the true parameter on average. The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram. The converse is true if the inflated frequency of the histogram is on the left hand side (i.e the average posterior over the SBC iterations overestimates prior).

            Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'. \citet{talts2020validating} describe this as under-dispersion of the posterior relative to the prior distribution on average. The average posterior is too narrow (or overconfident) relative to the spread of the prior resulting in inflated frequency on both ends of the rank statistic distribution.

@@

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/rhs.png}
                \caption{\textbf{Left:} Non uniform rank statistics with inflated frequency on the right side. \textbf{Right:} The average posterior distribution underestimates the prior distribution. The miscalibration from the rank statistic distribution comes from posterior samples underestimating the true parameter on average.}
                \label{fig:underestimation}
            \end{figure}

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/underdispersed.png}
                \caption{\textbf{Left}: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution). \textbf{Right}: Average posterior is underdispersed relative to the prior distribution. Non-uniformity in rank statistics comes from posteriors which are over-confident relative to the prior.}
                \label{fig:underdispersed}
            \end{figure}

@@

            \subsubsection{Effective sample size}
            Another univariate summary of a marginal posterior distribution is its effective sample size (ESS), which provides a measure of efficiency of the MCMC sampler. ESS estimates the number of effectively independent draws generated by a MCMC to estimate a target parameter. A poor ESS generally arises from high auto-correlation in the Markov chain, and thus is an indication of a highly dependent posterior sample. A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a sample from the target distribution when compared to a less efficient algorithm. If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), this is seen as evidence of greater efficiency.

            The ESS is denoted as $n_{\text{eff}}$ and is given by
            \begin{align}
                n_{\text{eff}} = \frac{NM}{1+2 \sum_{l=1}^L \hat{\rho}_l}.
            \end{align}
            When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag $l$ $\rho_l$ \citep{vehtari2021rank}. $L$ is a truncation lag chosen such that the sum of any two consecutive auto-correlation lags is negative \citep{geyer1992practical}. This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier auto-correlation estimates. Note that other approaches to estimating the true auto-correlation of the Markov chain and the ESS can be found in the literature. Details about this ESS formula, the calculation of the auto-correlation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.

@@

        The sampled momentum variable sets out the proposal path for $\theta$ governed by the Hamiltonian dynamics (represented by a set of differential equations). These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$. The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$. Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{n-1}, \alpha^{n-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$. A formal description of the HMC algorithm is provided in Appendix C.

        \subsubsection{Implementation}
        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study. Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model. The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}. Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.

@@

    \subsection{Sampling method 2: Off-set Mixture MCMC}
        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings \citep{metropolis1953equation, hastings1970monte} within Gibbs \citep{geman1984stochastic} algorithms and an implementation of the Kalman Filter \citep{kalman1960new} and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable. So a more recent simulation smoother is used from the statsmodels Python package by \citet{seabold2010statsmodels}. The simulation smoother software within the statsmodels package is based on \citet{durbin2012time}. Koopman who co-authored \citet{durbin2012time} also co-authored the software used by \citet{kim1998stochastic} which can be found at \citet{koopman1996ssfpack}.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian. Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring, adding an off-set, and taking the log of $y_t$, resulting in
        \begin{align}
        y_t^{*} &= \log(y_t^2 + c) = h_t + z_t, \label{eqn:transform}
        \end{align}
        where $z_t = \log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable. It is known that the error $z_t$ has mean -1.2704 and variance 4.93. The off-set $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}. By transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian. Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi-squared distribution. The mixture distribution is defined by
        \begin{align}
        f(z_t) = \sum_{i=1}^{I} q_if_N(z_t|m_i-1.2704, \nu_i^2),
        \end{align}
        where $K=7$, $f_N(z_t|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $i=1,2,\ldots, I=7$.  The values of these hyperparameters and component weights are applied in this research and reported in Appendix A.

@@

        The stochastic volatility model as described in section 2.1 follows a centered parameterisation. This describes the central location of the latent state distribution which is centered on the mean of log volatility $\mu$. A model can be reparameterised in a way that might improve MCMC performance. Reparameterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model. The reparameterisation available for a model is dependent on the MCMC approach. Some reparameterisation are available only to specific sampling strategies and may improve or impair a MCMC performance.

@@

        \subsubsection{Non-centered Gaussian mixture off-set model}
        The reparameterised model for the Gaussian mixture off-set is expressed differently due to the Kalman Filter. Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates. Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.

>>>>

\begin{abstract}
    Simulation Based Calibration (SBC) \citep{talts2020validating} is applied to analyse two commonly used, competing Markov chain Monte Carlo algorithms for estimating the posterior distribution of a stochastic volatility model. In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}. The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through the effective sample size. Results show that Stan's No-U-Turn sampler, an implementation of Hamiltonian Monte Carlo, produces a well-calibrated posterior estimate while the celebrated off-set mixture approach is less efficient and poorly calibrated, though model parameterisation also plays a role. Limitations and restrictions of generality are discussed.

@@

\section*{Acknowledgements}
First and foremost I would like to thank my supervisors Professor Catherine Forbes and Dr Lauren Kennedy. Your advice, guidance and patience has been invaluable. I am eternally grateful for the opportunity to learn from both of you.

I would also like to acknowledge the Masters research coordinator Professor Param Silvapulle for always checking in and keeping me and my fellow students on track.

Special thanks to Swen for her efforts in proofreading, providing feedback and being a reliable friend who I could talk with about my research.

And to my wife Dana who has supported and encouraged me from the very beginning. Thank you always for your love and patience every single day.

@@

    This research analyses and compares the calibration and efficiency of MCMC used to estimate stochastic volatility models using a simulation study. By calibration we refer to a MCMC algorithm that returns correct the posterior, on average, conditional on the correct model specification. The first algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}. The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the Kalman Filter and Metropolis Hastings algorithm. The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates on average.

    Comparisons of these algorithms are conducted using a methodology known as Simulation Based Calibration (SBC) \citep{talts2020validating}. Parameters are drawn from the prior distribution and used to create data sets from the generative stochastic volatility model. Posterior samples are then obtained from each of the MCMC methods conditional on the created data set. Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters given the data generating process. The key diagnostic metrics are the effective sample size, to measure the efficiency of the algorithm, and uniform rank statistics to determine calibration of the posterior estimates.

    One particular challenge to comparing models is the parameterisation. Different parameterisations of a model, whilst mathematically the same, can be very different in terms of sampling difficulty \citep{neal2003slice}. In this work, two different parameterisations of the stochastic volatility model are compared for each sampling method. An algorithm may be sensitive to different parameterisations of the same SV model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.

@@

\subsection{Challenge 1: Comparing posteriors with no ground truth}
    The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms. Convergence metrics assessed from the same MCMC output to be used for posterior inference are commonly used to check the performance of the MCMC chains. Most often, the effective sample size for each parameter is produced, as it indicates the number of approximately independent draws obtained from the dependent Markov chain for a target parameter \citep{gelman2013bayesian}. $\hat{R}$, the potential scale reduction factor or measure of variability between different chains \citep{gelman1992inference}, is also computed to gauge whether the MCMC algorithm has converged to the target posterior distribution.

@@

\subsection{Challenge 2: Limitations of a single simulation}
    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not observe the true model. A strategy around this is to evaluate a model and algorithm on simulated data. One approach is to simulate data from a generative model using known parameters. Then fit the same model on the simulated data and see if the true parameters can be recovered. This gives us the benefit of defining the true parameters of the data generating process to be estimated. If the model and algorithm cannot adequately capture the true parameter, then we cannot be confident that it will provide reliable estimates on real data.

@@

    Figure 2 shows marginal posterior distributions given a simulated data set of 1000 returns generated by the SV model. The true parameters are $\mu=-0.389$, $\sigma^2_{\eta}=0.0115$ and $\phi=0.547$, and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler, with 1000 post burn-in draws from the joint posterior distribution. The 95\% credible intervals for $\mu$ and $\sigma^2_{\eta}$ cover the true parameter. The true parameter for $\phi$ however, is in the tails and outside the interval. Such an analysis may incorrectly conclude that the model fails to adequately estimate the $\phi$ parameter. However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated data set.

@@

\subsection{Research Goal}
    The objective of this research is to evaluate the calibration of the HMC and KSC algorithms used to estimate SV models through a simulation study\footnote{The software used to produce all results and plots in this research are the R v4.2.3\citep{rlang}, Python v3.8.13 \citep{10.5555/1593511} and Stan v2.31.0 \citep{stan} programming languages, including open source R packages (tidyverse v1.3.2 \citep{tidyverse}, quantmod v0.4.20 \citep{quantmod}, cmdstanr 0.5.3 \citep{cmdstanr}, posterior v1.4.1 \citep{posteriorr}, bayesplot v1.4.10 \citep{bayesplot}, jsonlite v1.8.4 \citep{jsonlite}, arrow v12.0.1.1 \citep{arrow} and future.apply v1.11.0 \citep{RJ-2021-048}) and python packages (numpy v1.24.4 \citep{harris2020array}, scipy v.1.10.1 \citep{2020SciPy-NMeth}, statsmodels v.0.14.0 \citep{seabold2010statsmodels}, pandas v.2.0.3 \citep{mckinney-proc-scipy-2010} and pyarrow v12.0.1 \citep{arrow})}. As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations. To check the calibration of an algorithm, repeated independent simulations are required. This methodology is discussed in the next section.

@@

\section{Methodology}
This section introduces the simulation design used to evaluate whether a MCMC algorithm is returning the correct posterior estimates on average (that is, whether a MCMC algorithm calibrated). Section 3.1 discusses the procedure for running Simulation Based Calibration and Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.

    \subsection{Simulation Design}
        Simulation Based Calibration (SBC) checks the calibration of posterior estimates generated by MCMC algorithms. SBC is conducted by comparing the distribution of rank statistics to the uniform distribution which arises when an algorithm is correctly calibrated. For each $k$ of $K$ SBC iterations, one value for each parameter is drawn from the prior. Conditional on the drawn parameter values, a data set is generated from the model. Then a sampling method (originally using Stan in \citet{talts2020validating}) is used to obtain posterior samples given this data and the prior. The estimated posteriors are then compared to the true values by calculating the rank statistics. Under a calibrated algorithm, the $K$ rank statistics for each parameter should be uniformly distributed.

        To illustrate the SBC procedure, let $\theta$ be an arbitrary scalar parameter from a model and $y$ represent a data set. Start with a single parameter $\theta^{sim}$ drawn from the prior distribution
        \begin{align}
        \theta^{sim} \sim \pi(\theta),
        \end{align}
        then generate a data set conditional on $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, notation to condition on the model is suppressed.},
        \begin{align}
        y^{sim} \sim \pi (y|\theta^{sim}).
        \end{align}
        Take $B$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC)
        \begin{align}
        \{\theta_1,\dots , \theta_{B}\} \sim \pi (\theta | y^{sim}).
        \end{align}

        The last step is to calculate the rank statistic. The rank statistic for any one-dimensional function of parameters $f:\Theta\rightarrow\ \mathbb{R}$ is defined as
        \begin{align}
        r = \mathrm{rank}(\{f(\theta_1),\dots , f(\theta_{B})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}) < f(\theta^{sim})].
        \end{align}

        Equations (10)-(13) describes one iteration of SBC. To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.

        \citet{talts2020validating} prove that the sample of rank statistics for a given parameter follow a uniform distribution over the integers $\{0,1\dots,B\}$ if the MCMC algorithm is calibrated. Therefore, the rank statistics calculated from the SBC procedure are compared to the uniform distribution to determine if the algorithm is returning the correct posterior estimates on average.

        Another key result for calibration is that the posterior averaged over the data and true parameters is equal to the prior distribution. This is shown in the following expression
        \begin{align}
        \pi(\theta) &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})d\theta^{sim} dy^{sim} \\
        &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) d\theta^{sim} dy^{sim}.
        \end{align}
        That is, the average posterior for $\theta$ over the SBC iterations should approximately equal the prior distribution, if the algorithm is calibrated. Therefore, any deviation between the prior distribution and the average posterior draws over the SBC iterations may suggest that the sampling methodology is not producing the correct posterior on average.

@@

            The shape of the distribution of rank statistics gives insight into how the MCMC may be miscalibrated. Specifically, it gives information about how the average posterior deviates from the prior distribution of a given parameter. Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2020validating} for convenience. These display examples of posterior miscalibration and the corresponding non-uniform rank statistics.

            The left panel of Figure \ref{fig:underestimation} shows a distribution of rank statistics with inflated frequency on the right side of the histogram. This distribution is consistent with posterior samples that tend to underestimate the prior distribution on average, as shown on the right panel. The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram. The converse is true if the inflated frequency of the histogram is on the left hand side (i.e the average posterior over the SBC iterations overestimates prior).

            The left panel of Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'. \citet{talts2020validating} describe this as under-dispersion of the posterior relative to the prior distribution on average. The average posterior is too narrow (or overconfident) relative to the spread of the prior, as seen on the right panel, resulting in inflated frequency on both ends of the rank statistic distribution.

@@

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/rhs.png}
                \caption{\textbf{Left panel:} Non uniform rank statistics with inflated frequency on the right side. \textbf{Right panel:} The average posterior distribution underestimates the prior distribution. The miscalibration from the rank statistic distribution comes from posterior samples underestimating the prior distribution on average.}
                \label{fig:underestimation}
            \end{figure}

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/underdispersed.png}
                \caption{\textbf{Left panel}: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution). \textbf{Right panel}: Average posterior is underdispersed relative to the prior distribution. Non-uniformity in rank statistics comes from posteriors which are over-confident relative to the prior.}
                \label{fig:underdispersed}
            \end{figure}

@@

            \subsubsection{Effective sample size}
            Another univariate summary of a marginal posterior distribution is its effective sample size (ESS), which provides a measure of efficiency of the MCMC sampler. ESS estimates the number of effectively independent draws generated by a MCMC to estimate a target parameter. A poor ESS generally arises from high autocorrelation in the Markov chain, and thus is an indication of a highly dependent posterior sample. A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a sample from the target distribution when compared to a less efficient algorithm. If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), this is seen as evidence of greater efficiency.

            The ESS is denoted as $n_{\text{eff}}$ and is given by
            \begin{align}
                n_{\text{eff}} = \frac{NM}{1+2 \sum_{l=1}^L \hat{\rho}_l}.
            \end{align}
            When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the autocorrelation at lag $l$, $\rho_l$ \citep{vehtari2021rank}. $L$ is a truncation lag chosen such that the sum of any two consecutive autocorrelation lags is negative \citep{geyer1992practical}. This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier autocorrelation estimates. Note that other approaches to estimating the true autocorrelation of the Markov chain and the ESS can be found in the literature. Details about this ESS formula, autocorrelation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.

@@

        The sampled momentum variable sets out the proposal path for $\theta$ governed by the Hamiltonian dynamics (represented by a set of differential equations). These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$. The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$. Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{b-1}, \alpha^{b-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$. A formal description of the HMC algorithm is provided in Appendix C.

        \subsubsection{Implementation}
        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study. Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model. The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}. Stan samples the generative model directly and can flexibly handle complicated likelihood functions.

@@

    \subsection{Sampling method 2: Off-set Mixture MCMC}
        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings \citep{metropolis1953equation, hastings1970monte} within Gibbs \citep{geman1984stochastic} algorithms and an implementation of the Kalman Filter \citep{kalman1960new} and simulation smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable. So a more recent simulation smoother is used from the statsmodels Python package by \citet{seabold2010statsmodels}. The simulation smoother software within the statsmodels package is based on \citet{durbin2012time}. Koopman who co-authored \citet{durbin2012time} also co-authored the software used by \citet{kim1998stochastic} which can be found at \citet{koopman1996ssfpack}.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian. Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring, adding an off-set, and taking the log of $y_t$, resulting in
        \begin{align}
        y_t^{*} &= \log(y_t^2 + c) = h_t + z_t, \label{eqn:transform}
        \end{align}
        where $z_t = \log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable. It is known that the error $z_t$ has mean -1.2704 and variance 4.93. The off-set $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}. By transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian. Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi-squared distribution. The mixture distribution is defined by
        \begin{align}
        f(z_t) = \sum_{i=1}^{I} q_if_N(z_t|m_i-1.2704, \nu_i^2),
        \end{align}
        where $K=7$, $f_N(z_t|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $i=1,2,\ldots, I=7$.  The values of these hyperparameters and component weights are applied in this research and reported in Appendix A.

@@

        The stochastic volatility model as described in section 2.1 follows a centered parameterisation. This describes the central location of the latent state distribution which is centered on the mean of the log variance $\mu$. A model can be reparameterised in a way that might improve MCMC performance. Reparameterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model. The reparameterisation available for a model is dependent on the MCMC approach. Some reparameterisation are available only to specific sampling strategies and may improve or impair a MCMC performance.

@@

        \subsubsection{Non-centered Gaussian mixture off-set model}
        The reparameterised model for the Gaussian mixture off-set is expressed differently due to the Kalman Filter. Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates. Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log variance parameter $\mu$ entering the measurement equation.

        The following parameterisation is described as non centered in location. This follows the methodology outlined in \citet{strickland2008parameterisation}. Starting with the log chi-squared model
        \begin{align}
        y^{\ast}_t =& \space \, h_t + z_t \\
        h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t.
        \end{align}
        Reparameterise the state variable $h_t$ by subtracting the average log variance $\mu$. Let $g_t = \space \, h_t - \mu$. This leads to the state equation
        \begin{align}
        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}.
        \end{align}
        Return average log variance into measurement equation and rewrite as a function of the non centered state equation
        \begin{align}
        y^{\ast}_t =& \space \, g_t + \mu + z_t \\
        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t},
        \end{align}
        where the initial state is drawn from 
        \begin{align}
            g_1 \sim& \space \,\mathrm{N}\left(0, \frac{\sigma_{\eta}^2}{1-\phi^2}\right).
        \end{align}
        The mean of the log variance is now inside the measurement equation and demeaned from the state equation.
